Supporting Information for "Imperfect slope measurements drive overestimation in geometric models of lake depth"

### 2021-08-12 manuscript/figures.Rmd
move table S1 and Figure S6 to the main figures doc, #29

\captionsetup[figure]{name=Table}
\captionsetup[figure]{labelformat=empty}
```{r depth_model_metrics, fig.cap = "Table 1: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
# see tables/02_model_metrics.Rmd
knitr::include_graphics("../tables/02_model_metrics.pdf")
```
\captionsetup[figure]{name=Figure}
\captionsetup[figure]{labelformat=default}
\setcounter{figure}{2}

@@

>>>>

\captionsetup[figure]{name=Table}
\captionsetup[figure]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
```{r depth_model_metrics, fig.cap = "Table 1: Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset. The total number lakes is reported as n.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
# see tables/02_model_metrics.Rmd
knitr::include_graphics("../tables/01_predictors.pdf")
```

```{r depth_model_metrics, fig.cap = "Table 2: Model fit and predictive accuracy metrics (RMSE = root mean square error, R2 = coefficient of determination) for all combinations of true (in-lake slope, distance to the deepest point of the lake) and proxy (nearshore land slope, distance to lake center) metrics.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra='', fig.align='center'}
# see tables/02_model_metrics.Rmd
knitr::include_graphics("../tables/02_model_metrics.pdf")
```
\captionsetup[figure]{name=Figure}
\captionsetup[figure]{labelformat=default}
\setcounter{figure}{2}

@@

```{r depth_mode_importance, fig.cap="Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.", warning=FALSE, echo = FALSE, eval=TRUE, out.extra=''}
knitr::include_graphics("../figures/02_depth_model_importance-1")
```

### 2021-08-12 manuscript/si_template_2019.tex
move table S1 and Figure S6 to the main figures doc, #29

            \item \textbf{Figure S6:} Importance plot for random forest variables

            \item \textbf{Figure S7:} Characteristics of lakes with bathymetry versus non-bathymetry lakes

            \item \textbf{Figure S8:} Lake characteristics comparison among shape and reservoir classes

            \item \textbf{Figure S9:} Comparison between reported depth and bathymetry depth

            \item \textbf{Table S1:} Summary of lake characteristics for the present study

            \item \textbf{Table S2:} Effect of calculation technique on depth model fit and accuracy.

@@

\begin{figure}
      \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
      \caption{Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS7}
\end{figure}

@@

% source at tables/01_predictors.Rmd
\begin{table}
      \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset. The total number lakes is reported as n.} \label{tableS1}
      \centering
      \setlength\tabcolsep{1.5pt} % default value: 6pt
      \begin{tabular}{lllll}
            \hline
            Variable                            & Median      & Q25         & Q75          & n            \\
            \hline
            Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
            \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
            \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
            \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
            \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
            \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
            \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
            \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
            Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
            \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
            Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
            Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
            Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
            Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
            Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
            Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
            Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
            Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
            \hline
      \end{tabular}
\end{table}

>>>>

            \item \textbf{Figure S6:} Characteristics of lakes with bathymetry versus non-bathymetry lakes

            \item \textbf{Figure S7:} Lake characteristics comparison among shape and reservoir classes

            \item \textbf{Figure S8:} Comparison between reported depth and bathymetry depth

            \item \textbf{Table S1:} Effect of calculation technique on depth model fit and accuracy.

@@

% \begin{figure}
%       \noindent\includegraphics[width=0.85\textwidth]{../figures/02_depth_model_importance-1}
%       \caption{Importance plot for random forest variables showing increase in mean square error. Higher values indicate greater importance to model predictions. See Equation 1 for a definition of geometric max depth. HUC4 ID is a 'dummy' variable of geographic (hydrologic subbasin) location.}\label{figS7}
% \end{figure}

@@

% % source at tables/01_predictors.Rmd
% \begin{table}
%       \caption{Summary of lake characteristics for the present study (and for lakes in the contiguous United States). Predictor variables for computing random forest offsets (Equation 2) are printed in bold face. Dashes (-) indicate an identical sample size among this study and that of the contiguous United States from the National Hydrography Dataset. The total number lakes is reported as n.} \label{tableS1}
%       \centering
%       \setlength\tabcolsep{1.5pt} % default value: 6pt
%       \begin{tabular}{lllll}
%             \hline
%             Variable                            & Median      & Q25         & Q75          & n            \\
%             \hline
%             Max depth (m)                       & 8.2 (7)     & 4.6 (3.7)   & 14 (12)      & 4820 (17700) \\
%             \textbf{Area (ha)}                  & 55 (33)     & 21 (11)     & 140 (100)    & 4820 (17700) \\
%             \textbf{Island area (ha)}           & 0 (0)       & 0 (0)       & 0.18 (0.076) & 4820 (17700) \\
%             \textbf{Perimeter (m)}              & 4400 (3500) & 2500 (1800) & 8100 (7300)  & 4820 (17700) \\
%             \textbf{Shoreline development}      & 1.7 (1.7)   & 1.4 (1.4)   & 2.1 (2.2)    & 4820 (17700) \\
%             \textbf{Elevation (m)}              & 300 (340)   & 180 (210)   & 400 (460)    & 4820 (17700) \\
%             \textbf{Watershed-lake ratio}       & 7.8 (10)    & 3.8 (4.4)   & 17 (29)      & 4820 (17700) \\
%             \textbf{Deepest point distance (m)} & 180 (-)     & 110 (-)     & 290 (-)      & 4820 (-)     \\
%             Mean deepest point distance (m)     & 140 (-)     & 87 (-)      & 230 (-)      & 4820 (-)     \\
%             \textbf{Visual center distance (m)} & 240 (-)     & 160 (-)     & 390 (-)      & 4820 (-)     \\
%             Inlake slope (m/m)                  & 0.05 (-)    & 0.02 (-)    & 0.08 (-)     & 4820 (-)     \\
%             Inlake slope online (m/m)           & 0.06 (-)    & 0.03 (-)    & 0.14 (-)     & 4800 (-)     \\
%             Inlake slopes (m/m)                 & 0.06 (-)    & 0.03 (-)    & 0.1 (-)      & 4820 (-)     \\
%             Inlake slopes online (m/m)          & 0.07 (-)    & 0.03 (-)    & 0.15 (-)     & 4800 (-)     \\
%             Mean inlake slope (m/m)             & 0.04 (-)    & 0.02 (-)    & 0.09 (-)     & 4820 (-)     \\
%             Nearshore mean slope (m/m)          & 0.08 (-)    & 0.05 (-)    & 0.11 (-)     & 4820 (-)     \\
%             Nearshore slope online (m/m)        & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4590 (-)     \\
%             Nearshore slopes online (m/m)       & 0.08 (-)    & 0.04 (-)    & 0.13 (-)     & 4540 (-)     \\
%             \hline
%       \end{tabular}
% \end{table}
